A perfect storm and the natural endowments of trade-enabling infrastructure

### 2023-06-07 Slides/20min/20min.Rmd
Updated slides

#### Pre-event: 
- 1086-1208: Natural western channel closed because of gradual land rises (since the last ice age)
- 1672: Eastern Limfjord town Aalborg 2nd largest in DK 
- Other Limfjord towns lacked far behind under Aalborg's local hedgemony

@@

.pull-right[
![1838](http://www2.soundtoll.nl/sonttolregisters-333/Sonttolregisters-333_0055.jpg)
*Page from Sound Toll Register, (soundtoll.nl)*
]

@@

---
# The channel caused trade (1/2)

![Descriptive](Figures/Sound_toll.png)
---
name: str-results
# The channel caused trade (2/2)
.pull-left[
- **Data**: Panel of Danish ports 1750-1855 
- **Outcome**: Cargo ship traffic with origin or destination in a specified port
- PPML estimator following Silva & Tenreyro (2006) 
- In the order of 400% growth in trade in the region
- Local descriptive trade stats show the same story:
  + From 6 to 62 ships in Thisted harbour (1834-1876)
  + 997% increase in oats export, Thisted (1801-1853) 
]

.pull-right[
![STR_reg](Figures/STR_regression.PNG)
- *[Robustness: Only Origin/Destination](#onlyfromto)*<br>
- *[Port-level results](#str-results2)*
]

---
name: pop-ma
# Large population growth (1/2)
#### Market Access approach
![MA](Figures/wide_pop_ma.png)

@@

---
name: pop-dummy
# Large population growth (2/2)
#### Dummy approach
![Dummy](Figures/wide_pop_dummy.png)

???

---
name: mechanism1
.pull-left[
## Mechanism
- **First-order mechanism:** Trade (as documented)
- **Remaining questions**:

@@

  + *What carried the population growth?* 
  + **Answer:** Manufacturing and migration not fishing 
]

@@

**Fishing:**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/Wide_fish_new.png")
```

**Manufacturing:**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/Wide_manu_new.png")
```

@@

**Fishing:**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/Wide_wo_fish_dummy.png")
```

**Manufacturing:**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/Wide_wo_manu_dummy.png")
```

@@

**Data?** 
- Just use register data! 
- New method to construct panel sampling $P(Coin|t)_i$ based on Bayes and Monte Carlo
]

@@

---
name: arch
**Plot of effect to the likelihood of a coin finding being generated**
![Reg_arch](Figures/Arch_mod2.png)
[More details](#appendixarch)

---
name: other-results
# Conclusion 
.pull-left[
- **RQ:** How are we affected by geography? Does infrastructure cause development?
- Storm led to a channel and improved connectivity in 1834
- The population grew 22 percent in a generation in affected parishes 
- Population growth carried by trade, adaptation and rise of manufacturing
]

@@

---
name: str-results2
# The channel caused trade
![By port](Figures/Trade_effect_port_level_wide.png)

[APEs](#apes)

.footnote[
[Back](#str-results)
]

---
name: onlyfromto
# Traffic origin/destination
.pull-left[
### Only destination
![Dest](Figures/STR_reg_dest.png)
]

.pull-right[
### Only origin
![Origin](Figures/STR_reg_ori.PNG)
]

.footnote[
[Back](#str-results)
]

---
name: all-paramMA
### All feasible MA parameters
[Back](#pop-ma)

![Robust Pop MA](Figures/Varying_theta_alpha.png)

---
name: popma-outliers
### Robustness to outliers
[Back](#pop-ma)
![MAoutl](Figures/Robustness_to_outliers_pop.png)

---
name: specs-dummy
# All feasible specifications
![All_specs](Figures/All_specs_west.png)

.footnote[
[Back](#pop-dummy)
]

---
name: heterogendummy
# Urban/Rural heterogenous effects

![All_specs](Figures/all_data_FE_het_market_towns.png)

.footnote[
[Back](#pop-dummy)
]

@@

---
name: newspapers
# Newspapers
[Back](#MAadditional)

@@

---
name: apes
# APEs

![APE](Figures/Trade_APE_port_level_wide.png)

$$APE_h = exp(X\beta)\beta_h$$
.footnote[
[Back](#str-results2)
]

---
# Population growth - Poisson
#### Dummy

![Dummy](Figures/Possion.png)

.footnote[
[Back](#pop-dummy)
]

---
name: ma-pois
# Population growth 
#### Dummy Poisson

![Dummy](Figures/Possion_MA.png)

.footnote[
[Back](#pop-ma)
]

@@

**In equation form:**
- What we have: $P_i(t|c)$ 
- What we want: $P_i(c|t)$ 
]
--
.pull-right[
### The solution
Bayes formula and Monte Carlo!
- Bayes: $P_i(Coin|t) = \frac{P_i(t|Coin)P_i(Coin)}{P_i(t)}$ 
- Monte Carlo: For every coin:
  + Draw random number between $Year_{min}$ and $Year_{max}$ 
  + Repeat 1000 times 
  + Each of these represent a draw from the distribution we are interested in
  + Average to get estimate: $\widehat{P_i}(Coin|t)$
]

---
### Descriptive results
![Arch](Figures/Archeological_bootstrap_match.png)
*Error bars based on both sampling and dating uncertainty*
---
### Regression results
![Reg_arch](Figures/Arch_regs.png)

@@

---
#### Alternatives: All indicators

![Other_ind](Figures/Arch_ind_econ.png)

---
### All indicators 
![Other_ind](Figures/Arch_all.png)

---
# Breach $\rightarrow$ Migration
![Migr1](Figures/Wide_born_different_county.png)

>>>>

#### Pre-event: 
- 1085-1208: Natural western channel closed because of gradual land rises (since the last ice age)
- 1672: Eastern Limfjord town Aalborg 2nd largest in DK 
- Other Limfjord towns lacked far behind under Aalborg's local hedgemony

@@

.pull-right[
![1838](Figures/1734_STR_RUG.jpg)
*Page from Sound Toll Register in 1734 (www.rug.nl)*
]

@@

---
# The channel caused trade

![Ship traffic](Figures/Ship_trafic_channel.png)
*Descriptive statistics from Svalgaard (1977) sourced from customs archives*
---
# The channel caused trade

![Descriptive](Figures/Ship_trafic.png)
---
name: str-results
# The channel caused trade
.pull-left[
- **Data**: Panel of Danish ports 1750-1855 
- **Outcome**: Cargo ship traffic with origin or destination in a specified port
- PPML estimator following Silva & Tenreyro (2006) 
- In the order of 400% growth in trade in the region
- Local descriptive trade stats show the same story:
  + From 6 to 62 ships in Thisted harbour (1834-1876)
  + 997% increase in oats export, Thisted (1801-1853) 
]

.pull-right[
![STR_reg](Figures/STR_regressions.PNG)
*Larger on next page*
]

---
# The channel caused trade
![STR_reg](Figures/STR_regressions.PNG)

---
name: pop
# Population increase

.pull-left[
### Market Access approach
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/pop_ma.png")
```

**Robustness to:**
- [Comparison groups](#Rob_comp)
- [Parameter choices](#all-paramMA)
]
--
.pull-right[
### Dummy approach
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/pop_dummy.png")
```

**Robustness to:**
- [Comparison groups](#Rob_comp)
- [Pre-treatment characteristics](#pre_treat_adj)

@@

---
name: mechanism1
.pull-left[
# Mechanism
- **First-order mechanism:** Trade (as documented)
- **Remaining questions**:

@@

  + *What carried the population growth?* 
  + **Likely answer:** Fertility driven by rise of manufacturing and fishing
]

@@

**Fishing:**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/fish.png")
```

@@

**Manufacturing:**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/manu.png")
```

---
### But can this explain the population growth?

**Labor force as share of total pop:**
```{r echo=FALSE, out.width="70%", fig.align='center'}
knitr::include_graphics("Figures/all_occupations_share.png")
```

---
### Migration and fertility
.pull-left[
**People who live in different county than their birth county:**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/born_different.png")
```
]
--
.pull-right[
**Child/women ratio (ch. 1-5 years per f. 15-45):**
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/fertility.png")
```

@@

**Data?** 
- Just use register data! 
- New method to construct panel sampling $P(Coin|t)_i$ from dating range based on Bayes and Monte Carlo
]

@@

---
# Descriptive evidence
![Arch](Figures/Arch_descriptive.png)
---
name: arch
**Plot of effect to the likelihood of a coin finding being generated**
![Reg_arch](Figures/arch_dummy_coins.png)
[More details](#appendixarch)

---
name: arch
**Plot of effect to the likelihood of a building finding being generated**
![Reg_arch](Figures/arch_dummy_buildings.png)
[More details](#appendixarch)

---
name: other-results
# Conclusion 
.pull-left[
- **RQ:** How are we affected by geography? Does infrastructure cause development?
- Storm led to a channel and improved connectivity in 1834
- The population grew 22 percent in a generation in affected parishes 
- Population growth carried by trade and adaptation
- Intrinsic growth (fertility), fishing and manufacturing
]

@@

---
name: Rob_comp
## Robustness: Comparison groups

.pull-left[
#### MA approach
![MA_mult](Figures/Multiverse_MA.png)
]

.pull-right[
#### Dummy approach
![Dummy_mult](Figures/Multiverse_dummy.png)
]

[Back](#pop)

.footnote[
**Definitions:** A: <5km to coast; B: Copenhagen excl.; C: Control >100 km from Limfj.; D: <5km to Market town
]

---
name: all-paramMA
## Robustness: Parameter choices

```{r echo=FALSE, out.width="60%", fig.align='center'}
knitr::include_graphics("Figures/Multiverse_MA_param.png")
```

[Back](#pop)

---
name: pre_treat_adj
## Robustness: Pre-treat characteristics

.pull-left[
- Using the doubly robust estimator from Callaway & Sant’Anna (2021).
- Combines outcome regression with propensity score method. If either is valid, then the estimator is consistent.
- Only corrects for pre-treatment characteristics

**Columns:** (1): No adj.; (2): Adj. for age, occupation, fertility; (3): Adds log(Pop) as adj.
]

.pull-right[
![DR_est](Figures/CS_DR_estimate.png)
]

[Back](#pop)

@@

---
name: newspapers
# Newspapers

@@

@@

**In equation form:**
- What we have: $P_i(t|c)$: Probability of $t$ given one coin 
- What we want: $P_i(\{c\}|t)$: Probability of *any* coins in time $t$
]
--
.pull-right[
### The solution
Bayes formula and Monte Carlo!
- Simple to sample from this distribution
- Sample $t$ from each coin 
- Count frequency of coin findings
]

$$P_i(\{findings\}|t)=\left[1-\prod_{c=1}^{K_i} \left( 1 - P(t|finding_c) \right)\right] P(\{findings\})$$
---
### All indicators 
![Other_ind](Figures/Arch_all.png)

@@
